### 2020-07-10 submission/manuscript.tex

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\begin{document}

\vspace{35mm}

\section{\texorpdfstring{The initial gut microbiota and response to
antibiotic perturbation influence \emph{Clostridioides difficile}
colonization in
mice}{The initial gut microbiota and response to antibiotic perturbation influence Clostridioides difficile colonization in mice}}\label{the-initial-gut-microbiota-and-response-to-antibiotic-perturbation-influence-clostridioides-difficile-colonization-in-mice}

\vspace{35mm}

Sarah Tomkovich\({^1}\), Joshua M.A.~Stough\({^1}\), Lucas
Bishop\({^1}\), Patrick D. Schloss\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

\(1\) Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

The gut microbiota has a key role in determining susceptibility to
\emph{Clostridioides difficile} infections (CDIs). However, much of the
mechanistic work examining CDIs in mouse models use animals obtained
from a single university colony or vendor. We treated mice from 6
different sources (2 University of Michigan colonies and 4 vendors) with
a single clindamycin dose, followed by a \emph{C. difficile} challenge 1
day later and then measured \emph{C. difficile} colonization levels
through 9 days post-infection. The microbiota was profiled via 16S rRNA
gene sequencing to examine the variation across sources and alterations
due to clindamycin treatment and \emph{C. difficile} challenge. While
all sources of mice were colonized 1-day post-infection, variation
emerged from days 3-7 post-infection with animals from some sources
colonized with \emph{C. difficile} for longer and at higher levels. We
identified bacteria that varied in relative abundance across sources and
throughout the experiment. Some bacteria were consistently impacted by
clindamycin treatment in all sources of mice including
\emph{Lachnospiraceae}, \emph{Ruminococcaceae}, and
\emph{Enterobacteriaceae}. To identify bacteria that were most important
to colonization regardless of the source, we created logistic regression
models that successfully classified mice based on whether they cleared
\emph{C. difficile} by 7 days post-infection using baseline,
post-clindamycin, and post-infection community composition data. With
these models, we identified 4 bacteria that varied across sources
(\emph{Bacteroides}), were altered by clindamycin
(\emph{Porphyromonadaceae}), or both (\emph{Enterobacteriaceae} and
\emph{Enterococcus}). Microbiota variation across sources better
emulates human interindividual variation and can help identify bacterial
drivers of phenotypic variation in the context of CDIs.

\subsection{Importance}\label{importance}

\emph{Clostridioides difficile} is a leading nosocomial infection.
Although perturbation to the gut microbiota has been established as a
key risk factor, there is variation in who becomes asymptomatically
colonized, develops an infection, or has an infection with adverse
outcomes. \emph{C. difficile} infection (CDI) mouse models are widely
used to answer a variety of \emph{C. difficile} pathogenesis questions.
However, the inter-individual variation between mice is less than what
is observed in humans, particularly if just one source of mice is used.
In this study, we administered clindamycin to mice from 6 different
breeding colonies and challenged them with \emph{C. difficile}.
Interestingly, only a subset of the bacteria that vary across sources
were associated with how long \emph{C. difficile} was able to colonize.
Future studies examining the interplay between the microbiota and
\emph{C. difficile} should consider using mice from multiple sources to
better reflect human interindividual variation.

\newpage

\subsection{Introduction}\label{introduction}

Antibiotics are a common risk factor for \emph{Clostridioides difficile}
infections (CDIs), but there is variation in who goes on to develop
severe or recurrent CDIs after exposure (1, 2). Additionally,
asymptomatic colonization, where \emph{C. difficile} is detectable, but
symptoms are absent has been documented in infants and adults (3, 4).
The intestinal microbiome has been implicated in asymptomatic
colonization (5, 6), susceptibility to CDIs (7), and adverse CDI
outcomes (9--12).

Mouse models of CDIs have been a great tool for understanding \emph{C.
difficile} pathogenesis (13). The number of CDI mouse model studies has
grown substantially since Chen et al. published their C57BL/6 model in
2008, which disrupted the gut microbiota with antibiotics to enable
\emph{C. difficile} colonization and symptoms such as diarrhea and
weight loss (14). CDI mouse models have been used to examine
translationally relevant questions regarding \emph{C. difficile},
including the role of the microbiota and efficacy of potential
therapeutics for treating CDIs (15). However, variation in the
microbiome between mice from the same breeding colony is much less than
the variation observed between humans (16, 17). Additionally, studying
the contribution of the microbiota to a particular disease phenotype in
one set of lab mice after the same perturbation could yield a number of
findings of which only a fraction may be driving the phenotype.

In the past, our group has attempted to introduce more microbiome
variation into the CDI mouse model by using a variety of antibiotic
treatments (18--21). An alternative approach to maximize microbiome
variation is to use mice from multiple sources (22, 23). Microbiome
differences between different mouse vendors have been well documented
and shown to influence susceptibility to a variety of diseases (24, 25),
including enteric infections (22, 23, 26--30). Additionally, research
groups have observed different CDI outcomes in mice despite using
similar models and the microbiome has been proposed as one factor
potentially mediating CDI susceptibility and outcomes (13, 18, 21,
31--33). Here we examined how variations in the baseline microbiome and
responses to clindamycin treatment in C57BL/6 mice from six different
sources influenced susceptibility to \emph{C. difficile} colonization
and the time needed to clear the infection.

\subsection{Results}\label{results}

\textbf{Microbiota variation is high between mice from different
sources.} To test how the microbiotas of mice from different sources
impact colonization dynamics after clindamycin exposure, we utilized
C57BL/6 mice from 6 different sources: two colonies from the University
of Michigan that were split from each other 10 years ago (the Young and
Schloss lab colonies) and four commercial source: the Jackson
Laboratory, Charles River Laboratories, Taconic Biosciences, and Envigo
(which was formerly Harlan). These 4 vendors were chosen because they
represent commonly used vendors for CDI studies in mice (26, 34--40). We
used 16S rRNA gene sequencing to characterize the fecal bacterial
communities across sources at baseline.

Since antibiotics and other risk factors of CDIs are associated with
decreased microbiota diversity (41), we first examined alpha diversity
across the 6 sources of mice. Intially, there was a significant
difference in the number of observed operational taxonomic units (OTUs),
but not Shannon diversity index (\emph{P}\textsubscript{FDR} = 0.03 and
\emph{P}\textsubscript{FDR} = 0.052, respectively) across sources of
mice (Fig. 1A-B and Table S1). Next, we compared the community structure
of mice from the 6 sources over the course of the experiment using
principal coordinate analysis (PCoA) of the \(\theta_{YC}\) distances.
Permutational multivariate analysis of variance (PERMANOVA) analysis
revealed source and cage interactions explained most of the observed
variation across fecal communities explained most of the observed
community variation (combined R\textsuperscript{2} = 0.90, \emph{P} =
0.0001, Fig. 3 and Table S2). Mice that are co-housed have been shown to
have similar gut microbiotas due to coprophagy (42) and since mice
within the same source were housed together, it is not surprising that
cage also contributed to the observed microbiota variation. Since we
conducted two experiments approximately 3 months apart, we also compared
baseline communities across the 2 experiments. Experiment and cage
significantly explained the observed variation only for the Schloss and
Young lab mouse colonies (Fig. S2A-B and Table S2). However, most of the
vendors also clustered by experiment (Fig. S2C-D, F), suggesting there
was some community variation between the 2 experiments within each
source, particularly for Schloss, Young, and Envigo mice (Fig. S2G-H).
After finding differences at the community level, we next identified the
bacteria that initially varied across sources of mice. There were 268
OTUs with relative abundances that varied across sources (Fig. 1D and
Table S3). By using mice from six sources we were able to increase the
number of microbiota communities to test with the clindamycin \emph{C.
difficile} colonization mouse model.

\textbf{Clindamycin treatment renders all mice susceptible to \emph{C.
difficile} 630 colonization, but clearance time varies across sources.}
After the acclimation period, all mice were treated with 10 mg/kg
clindamycin via intraperitoneal injection and one day later challenged
with 10\textsuperscript{3} \emph{C. difficile} 630 spores (Fig. 2A).
Clindamycin was chosen because we have previously demonstrated mice are
rendered susceptible, but consistently cleared \emph{C. difficile}
within 9 days (21, 43), clindamycin is frequently implicated with human
CDIs (44), and is also part of the antibiotic treatment for the
frequently cited 2008 CDI mouse model (14). The day after infection,
\emph{C. difficile} was detectable in all mice at a similar level
(median CFU range: 2.2e+07-1.3e+08; \emph{P}\textsubscript{FDR} = 0.15),
indicating clindamycin rendered all mice susceptible regardless of
source (Fig. 2B). Interestingly, variation in \emph{C. difficile} CFU
levels across sources of mice emerged from days 3-7 post-infection (all
\emph{P}\textsubscript{FDR} \(\le\) 0.019; Fig. 2B and Table S4),
suggesting mouse source is associated with \emph{C. difficile}
clearance. We conducted two experiments and while the colonization
dynamics were similar across most sources of mice, there was some
variation between the 2 experiments, particularly for the Schloss and
Envigo mice (Fig. S2A-B). Although \emph{C. difficile} 630 causes mild
symptoms in mice comparied to other \emph{C. difficile} strains (45), we
also saw that weight change significantly varied across sources of mice
with the most weight lost two days post-infection (Fig. 2C and Table
S5). Importantly, there was also one Jackson and one Envigo mouse that
died between 1- and 3-days post-infection during the second experiment.
Interestingly, mice obtained from Jackson, Taconic, and Envigo tended to
lose more weight, have higher \emph{C. difficile} CFU levels and take
longer to clear the infection compared to the other sources of mice
(although there was variation between experiments with Schloss and
Envigo mice), which was particularly evident 7 days post-infection (Fig.
2B-C, Fig. S2C-D), when 57\% of the mice were still colonized with
\emph{C. difficile} (Fig. S2E). By 9 days post-infection the majority of
the mice from all sources had cleared \emph{C. difficile} (Fig. 2C) with
the exception of 1 Taconic mouse from the first experiment and 2 Envigo
mice from the second experiment. Thus, clindamycin rendered all mice
susceptible to \emph{C. difficile} 630 colonization, regardless of
source, but there was significant variation in disease phenotype across
the sources of mice.

\textbf{Clindamycin treatment alters bacteria in all sources, but a
subset of bacterial differences across sources persists.} Given the well
known variation in mouse microbiomes across breeding colonies (25), we
hypothesized that variation in \emph{C. difficile} clearance would be
explained by microbiota variation across the 6 sources of mice. We used
16S rRNA sequencing to characterize microbiota variation across sources
after clindamycin treatment and throughout the rest of the experiment.
PERMANOVA analysis revealed source was the dominant factor that
explained the observed variation across fecal communities
(R\textsuperscript{2} = 0.35, \emph{P} = 0.0001) followed by
interactions between cage (mice from the same sources were housed
together, primarily at a density of 2 mice per cage) and day of the
experiment (Movie S1 and Table S2). Since the majority of the
perturbations happened over the initial days of the experiment, we
decided to focus on the communities after clindamycin treatment (day 0)
and post-infection (day 1).

As expected, clindamycin treatment decreased richness and Shannon
diversity across all sources of mice (Fig. 3A-B). Interestingly,
significant differences in diversity metrics across sources
(\emph{P}\textsubscript{FDR} \textless{} 0.05) emerged after clindamycin
treatment, with Charles River mice having higher richness and Shannon
diversity than most of the other sources (Fig 3A-B and Table S1). The
variation in community structures was also diminished by clindamycin
treatment (Fig. 3C and Table S2). Next we identified the bacteria that
shifted after clindamycin treatment, regardless of source by analyzing
all mice that had sequence data from fecal samples collected at baseline
and after clindamycin treatment. We identified 153 OTUs that were
altered after clindamycin treatment (Fig. 3D and Table S6) in most mice.
Despite the significant impact of clindamycin, there were still 18 OTUs
(Fig. 3E and Table S3) with relative abundances that significantly
varied across sources. Interestingly, when we compared the list of
significant clindamycin impacted bacteria with the bacteria that varied
across sources post-clindamycin, we found 4 OTUs
(\emph{Enterobacteriaceae (OTU 1)}, \emph{Lachnospiraceae} (OTU 130),
\emph{Lactobacillus} (OTU 6), \emph{Enterococcus} (OTU 23)) overlapped
(Fig. 3D-E and Tables S3, S6). These findings demonstrate that
clindamycin has a consistent impact on the fecal bacterial communities
of mice from all sources and only a subset of the OTUs continued to vary
across sources.

\textbf{Microbiota variation across sources is maintained after \emph{C.
difficile} challenge.} Post-infection, significant differences in
diversity metrics were maintained across sources
(\emph{P}\textsubscript{FDR} \textless{} 0.05, Fig 4A-B and Table S1).
Although the Charles River mice had more diverse microbiotas, the Young
and Schloss lab mice were also able to clear \emph{C. difficile} faster
than some of the other sources, suggesting microbiota diversity alone
does not explain the observed variation in \emph{C. difficile}
colonization across sources. Source and cage interaction continued to
explain most of the observed community variation (combined
R\textsuperscript{2} = 0.88, respectively; \emph{P} = 0.0001; Fig. 4D
and Table S2). After \emph{C. difficile} challenge, there were 44 OTUs
(Fig. 4D and Table S3) with significantly different relative abundances
across sources. In spite of the experimental perturbations that
occurred, there were 12 OTUs with relative abundances that consistently
varied across sources at baseline, post-clindamycin, and post-infection.
Importantly, some of the OTUs that consistently varied across sources
also shifted with clindamycin treatment. For example, \emph{Proteus}
increased after clindamycin treatment, but only in Taconic mice.
\emph{Enterococcus} was primarily found only in mice purchased from
commercial vendors and also increased after clindamycin treatment. In
summary, mouse bacterial communities varied significantly according to
source throughout the course of the experiment and a consistent subset
of bacterial taxa remained different across sources regardless of
clindamycin and \emph{C. difficile} challenge.

\textbf{Baseline, post-clindamycin, and post-infection community data
can predict mice that will clear \emph{C. difficile} by 7 days
post-infection.} After identifying taxa that varied by source, changed
after clindamycin treatment, or both, we next wanted to determine which
taxa were influencing the variation in \emph{C. difficile} colonization
at day 7 (Fig. 2D, Fig. S2C). We trained three L2-regularized logistic
regression models with input bacterial community data from the baseline,
post-clindamycin, and post-infection timepoints of the experiment to
predict \emph{C. difficile} colonization status on day 7 (Fig. S3A-B).
All models were better at predicting \emph{C. difficile} colonization
status on day 7 than random chance (all \emph{P} \(\le\) 1.6e-17; Table
S7). Interestingly, the model based on the post-clindamycin (day 0)
community OTU data performed significantly better than the other models
with an area under the receiving operator characteristic curve (AUROC)
of 0.78 (\emph{P}\textsubscript{FDR} \(\le\) 3.3e-06 for pairwise
comparisons; Table S7). Thus, we were able to use bacterial relative
abundance data alone to differentiate mice that had cleared \emph{C.
difficile} before day 7 from the mice still colonized with \emph{C.
difficile} at that timepoint. Interestingly, the model built with OTU
relative abundance data post-clindamycin treatment had the best
performance, suggesting how the bacterial community responds to
clindamycin treatment has the greatest influence on subsequent \emph{C.
difficile} colonization dynamics.

Next, to examine the bacteria that were driving each model's
performance, we pulled out the top 20 OTUs that had the highest absolute
feature weights in each of the 6 models (Table S8). First, we looked at
OTUs from the model with the best performance that was based on the
post-clindamycin treatment bacterial community data. While most of the
20 OTUs had low relative abundances on day 0, \emph{Enterobacteriace},
\emph{Bacteroides} and \emph{Proteus} had high relative abundances in at
least one source of mice and significantly varied across sources (Fig.
5A). Next, the top 20 taxa from each model were compared to the list of
taxa that varied across source at the same timepoint (Fig. 1D, 3E, 4D
and Table S3) and the taxa that were altered by clindamycin treatment
(Fig. 3D and Table S6). We found a subset of OTUs that were important to
the model and overlapped with bacteria that varied by either source,
clindamycin treatment, or both (Fig. S4). Combining the overall results
for the 3 models identified 14 OTUs associated with source, 21 OTUs
associated with clindamycin treatment, and 6 OTUs associated with both
(Fig. 5B). Several OTUs (\emph{Bacteroides (OTU 2), Enterococcus (OTU
23), Enterobacteriaceae (OTU 1), Porphyromonadaceae (OTU 7)}) appeared
across at least 2 models, so we examined how the relative abundances of
these OTUs varied over the course of the experiment (Fig. 6). Throughout
the experiment, there was at least 1 timepoint where relative abundances
of these OTUs significantly varied across sources (Table S9).
Interestingly, there were no OTUs that emerged as consistently enriched
or depleted in mice that were colonized past 7 days post-infection with
\emph{C. difficile} 630, suggesting multiple bacteria influence the time
needed to clear the infection. Together, these results suggest the
initial bacterial communities and their responses to clindamycin
influence the clearance of \emph{C. difficile}.

\subsection{Discussion}\label{discussion}

By running our CDI model with mice from 6 different sources, we were
able to identify bacterial taxa that were unique to sources throughout
the experiment as well as taxa that were universally impacted by
clindamycin. We trained L2 logistic regression models with baseline,
post-clindamycin treatment, and post-infection fecal community data that
could predict whether mice cleared \emph{C. difficile} by 7 days
post-infection better than random chance. We identified
\emph{Bacteroides (OTU 2), Enterococcus (OTU 23), Enterobacteriaceae
(OTU 1), Porphyromonadaceae (OTU 7)} (Fig. 6) as candidate bacteria
within these communities that were influencing variation in \emph{C.
difficile} colonization dynamics since these bacteria were all important
in the logistic regression models and varied by source, were impacted by
clindamycin treatment, or both. Overall, our results demonstrate
clindamycin is sufficient to render mice from multiple sources
susceptible to CDI and only a subset of the interindividual microbiota
variation across mice from different sources was associated with the
time needed to clear \emph{C. difficile}.

Other studies have taken similar approaches by using mice from multiple
sources to identify bacteria that either promote colonization resistance
or increase susceptibility to enteric infections (22, 23, 26--30). For
example, in the context of \emph{Salmonella} infections,
\emph{Enterobacteriaceae} and segmented filamentous bacteria have
emerged as protective (22, 27). A previous study with \emph{C.
difficile} identified an endogenous protective \emph{C. difficile}
strain LEM1 that bloomed after antibiotic treatment in mice from Jackson
or Charles River Laboratories, but not Taconic that protected mice
against the more toxigenic \emph{C. difficile} VPI10463 (26). Given that
we obtained mice from the same vendors, we checked all mice for
endogenous \emph{C. difficile} by plating stool samples that were
collected after clindamycin treatment. However, we did not identify any
endogenous \emph{C. difficile} strains prior to challenge, suggesting
there were no endogenous protective strains in the mice we received and
other bacterial taxa mediated the variation in \emph{C. difficile}
colonization across sources.

Although all mice were susceptible to \emph{C. difficile} colonization,
mice from Jackson, Taconic, and Envigo mice tended to remain colonized
through at least 7 days post-infection. We identified a subset of
bacteria that were important in predicting whether a mouse was still
colonized with \emph{C. difficile} 7 days post-infection. These results
suggest a subset of the bacterial community is responsible for
determining the length of time needed to clear \emph{C. difficile}
colonization.

Differences in CDI mouse model studies have been attributed to different
intestinal microbiotas of mice from different sources. For example,
groups using the same clindamycin treatment and C57BL/6 mice had
different \emph{C. difficile} outcomes, one having sustained
colonization (32), while the other had transient (18), despite both
using \emph{C. difficile} VPI 10643. Baseline differences in the
microbiota composition have been hypothesized to partially explain the
differences in colonization outcomes and overall susceptibility to
\emph{C. difficile} after treatment with the same antibiotic (13, 31).
The bacterial perturbations induced by clindamycin treatment have been
well characterized and our findings agree with previous CDI mouse model
work demonstrating \emph{Enterococcus} and \emph{Enterobacteriaceae}
were associated with \emph{C. difficile} susceptibility and
\emph{Porpyhromonadaceae}, \emph{Lachnospiraceae},
\emph{Ruminococcaceae}, and \emph{Turicibacter} were associated with
resistance (19, 21, 32, 33, 43, 46--48). While we have demonstrated that
susceptibility is uniform across sources of mice after clindamycin
treatment, there could be different outcomes for either susceptibility
or clearance in the case of other antibiotic treatments. The \emph{C.
difficile} strain used could also be contributing to the variation in
\emph{C. difficile} outcomes seen across different research groups. For
example, a group found differential colonization outcomes after
clindamycin treatment, with \emph{C. difficile} 630 and M68 eventually
becoming undectable while BI-7 remained detectable up to 70 days
post-treatment (47). We found the time needed to naturally clear
\emph{C. difficile} varied across sources of mice implying that at least
in the context of the same perturbation, microbiota differences seemed
to influence infection outcome more than susceptibility. More
importantly, we were able to reduce the variation observed across
sources to identify a subset of OTUs that were also important for
predicting \emph{C. difficile} colonization status 7 days
post-infection. Since all but 3 mice eventually cleared \emph{C.
difficile} 630 by 9 days post-infection and the model built with the
post-clindamycin OTU relative abundance data had the best performance,
our results suggest clindamycin treatment had a large role in
determining \emph{C. difficile} susceptibility and clearance than the
source of the mice.

Our approach successfully increased the diversity of murine bacterial
communities tested in our clindamycin \emph{C. difficile} model. One
alternative approach that has been used in some CDI studies (49--54) is
to associate mice with human microbiotas. However, a major caveat to
this method is the substantial loss of human microbiota community
members upon transfer to mice (55, 56). Additionally with the exception
of 2 recent studies (49, 50), most of the CDI mouse model studies to
date associated mice with just 1 types of human microbiota either from a
single donor or a single pool from multiple donors (51--54), which does
not aid in the goal of modeling the interpersonal variation seen in
humans to understand how the microbiota influences susceptiblity to CDIs
and adverse outcomes. Importantly, our study using mice from 6 different
sources increased the variation between groups of mice compared to using
1 source alone, to better reflect the inter-individual microbiota
variation observed in humans. Encouragingly, decreased
\emph{Bifidobacterium}, \emph{Porphyromonas}, \emph{Ruminococcaceae} and
\emph{Lachnospiraceae} and increased \emph{Enterobacteriaceae},
\emph{Enterococcus}, \emph{Lactobacillus}, and \emph{Proteus} have all
been associated with human CDIs (7) and were well represented in our
study, suggesting most of the mouse sources are suitable for gaining
insights into microbiota associated factors influencing \emph{C.
difficile} colonization and infections in humans. An important exception
was \emph{Enterococcus}, which was primarily absent from the mice from
University of Michigan colonies and \emph{Proteus}, which was only found
in Taconic mice. Importantly, the fact that some CDI-associated bacteria
were only found in a subset of mice has important implications for
future CDI mouse model studies.

There are several limitations to our work. The microbiome is composed of
viruses, fungi, and parasites in addition to bacteria, and these
non-bacterial members can also vary across mouse sources (57, 58). While
our study focused solely on the bacterial portion, viruses and fungi
have also begun to be implicated in the context of CDIs or FMT
treatments for recurrent CDIs (35, 59--62). Beyond community
composition, the metabolic function of the microbiota also has a CDI
signature (20, 48, 63, 64) and can vary across mice from different
sources (65). For example, microbial metabolites, particularly secondary
bile acids and butyrate production, have been implicated as important
contributors to \emph{C. difficile} resistance (33, 47). Although, we
only looked at composition, \emph{Ruminococcaceae} and
\emph{Lachnospiraceae} both emerged as important taxa for classifying
day 7 \emph{C. difficile} colonization status and metagenomes from these
bacteria have been shown to contain the bile acid-inducible gene cluster
necessary for secondary bile acid formation and ability to produce
butyrate (52, 66). Interestingly, butyrate has previously been shown to
vary across vendors and mediated resistance to \emph{Citrobacter
rodentium} infection, a model of enterohemorrhagic and enteropathogenic
\emph{Escherichia coli} infections (23). Evidence for immunological
toning differences in IgA and Th17 cells across mice from different
vendors have also been documented and (67, 68) may also influence
response to CDI, particularly in the context of severe CDIs (69, 70).
The outcome after \emph{C. difficile} exposure depends on a multitude of
factors, including age, diet, and immunity; all of which are also
influenced by the microbiota.

We have demonstrated that the ways baseline microbiotas from different
mouse sources respond to clindamycin treatment influences the length of
time mice remained colonized with \emph{C. difficile} 630. For those
interested in dissecting the contribution of the microbiome to \emph{C.
difficile} pathogenesis and treatments, using multiple sources of mice
may yield more insights than a single model alone. Furthermore, for
studies wanting to examine the interplay between a particular bacterial
taxon such as \emph{Enterococcus} and \emph{C. difficile}, these results
could serve as a resource for selecting which mice to order to address
the question. Using mice from multiple sources helps model the
interpersonal microbiota variation among humans to aid our understanding
of how the gut microbiota contributes to CDIs.

\newpage

\subsection{Acknowledgements}\label{acknowledgements}

This work was supported by the National Institutes of Health
(U01AI124255). ST was supported by the Michigan Institute for Clincial
and Health Research Postdoctoral Translation Scholars Program
(UL1TR002240). We thank members of the Schloss lab for feedback on
planning the experiments and data presentation, as well as code
tutorials and feedback through Code club. In particular, we want to
thank Begüm Topçuoğlu for help with implementing L2 logistic regression
models using her
\href{https://github.com/SchlossLab/ML_pipeline_microbiome}{pipeline},
Ana Taylor for help with media preparation and sample collection, and
Nicholas Lesniak for his critical feedback on the manuscript. We also
thank members of Vincent Young's lab, particularly Kimberly Vendrov, for
guidance with the \emph{C. difficile} infection mouse model and donating
the mice. We also want to thank the Unit for Laboratory Animal Medicine
at the University of Michigan for maintaining our mouse colony and
providing the institutional support for our mouse experiments. Finally,
we thank Kwi Kim, Austin Campbell, and Kimberly Vendrov for their help
in maintaining the Schloss lab's anaerobic chamber.

\newpage

\subsection{Materials and Methods}\label{materials-and-methods}

\textbf{(i) Animals.} All experiments were approved by the University of
Michigan Animal Care and Use Committee (IACUC) under protocol number
PRO00006983. Female C57BL/7 mice were obtained from 6 different sources:
The Jackson Laboratory, Charles River Laboratories, Taconic Biosciences,
Envigo, and two colonies at the University of Michigan (the Schloss lab
colony and the Young lab colony). The Young lab colony was originally
established with mice purchased from Jackson, and the Schloss lab colony
was established 10 years ago with mice donated from the Young lab. The 4
groups of mice purchased from vendors were allowed to acclimate to the
University of Michigan mouse facility for 13 days prior to starting the
experiment. At least 4 female mice (age 5-10 weeks) were obtained per
source and mice from the same source were primarily housed at a density
of 2 mice per cage. The experiment was repeated once, approximately 3
months after the start of the first experiment.

\textbf{(ii) Antibiotic treatment.} After the 13-day acclimation period
and 1 day prior to challenge (Fig. 1A), all mice received 10 mg/kg
clindamycin (filter sterilized through a 0.22 micron syringe filter
prior to administration) via intraperitoneal injection.

\textbf{(iii) \emph{C. difficile} infection model.} Mice were challenged
with 10\textsuperscript{3} spores of \emph{C. difficile} strain 630 via
oral gavage post-infection 1 day after clindamycin treatment as
described previously (21). Mice weights and stool samples were taken
daily through 9 days post-challenge. Collected stool was split for
\emph{C. difficile} CFU quantification and 16S rRNA sequencing analysis.
\emph{C. difficile} quantification stool samples were transferred to the
anaerobic chamber, serially diluted in PBS, plated on
taurocholate-cycloserine-cefoxitin-fructose agar (TCCFA) plates, and
counted after 24 hours of incubation at 37°C under anaerobic conditions.
A sample from the day 0 timepoint (post-clindamycin and prior to
\emph{C. difficile} challenge) was also plated on TCCFA to ensure mice
were not already colonized with \emph{C. difficile} prior to infection.
There were 3 deaths recorded over the course of the experiment, 1
Taconic mouse died prior to \emph{C. difficile} challenge and 1 Jackson
and 1 Envigo mouse died between 1- and 3-days post-infection. Mice were
categorized as cleared when no \emph{C. difficile} was detected in the
first serial dilution (limit of detection: 100 CFU). Stool samples for
16S rRNA sequencing were snap frozen in liquid nitrogen and stored at
-80°C until DNA extraction.

\textbf{(iv) 16S rRNA sequencing.} DNA was extracted from -80°C stored
stool samples using the DNeasy Powersoil HTP 96 kit (Qiagen) and an
EpMotion 5075 automated pipetting system (Eppendorf). The V4 region was
amplified for 16S rRNA with the AccuPrime Pfx DNA polymerase (Thermo
Fisher Scientific) using custom barcoded primers, as previously
described (71). The ZymoBIOMICS microbial community DNA standards was
used as a mock community control (72) and water was used as a negative
control per 96-well extraction plate. The PCR amplicons were cleaned up
and normalized with the SequalPrep normalization plate kit (Thermo
Fisher Scientific). Amplicons were pooled and quantified with the KAPA
library quantification kit (KAPA biosystems), prior to sequencing using
the MiSeq system (Illumina).

\textbf{(v) 16S rRNA gene sequence analysis.} mothur (v. 1.43) was used
to process all sequences (73) with a previously published protocol (71).
Reads were combined and aligned with the SILVA reference database (74).
Chimeras were removed with the VSEARCH algorithm and taxonomic
assignment was completed with a modified version (v16) of the Ribosomal
Database Project reference database (v11.5) (75) with an 80\% cutoff.
Operational taxonomic units (OTUs) were assigned with a 97\% similarity
threshold using the opticlust algorithm (76). To account for uneven
sequencing across samples, samples were rarefied to 5,437 sequences
1,000 times for alpha and beta diversity analyses. PCoAs were generated
based on \(\theta_{YC}\) distances. Permutational multivariate analysis
of variance (PERMANOVA) was performed on mothur-generated
\(\theta_{YC}\) distance matrices with the adonis function in the vegan
package (77) in R (78).

\textbf{(vi) Classification model training and evaluation.} Models were
generated based on mice that were categorized as either cleared or
colonized 7 days post-infection and had sequencing data from the
baseline (day -1), post-clindamycin (day 0), and post-infection (day 1)
timepoints of the experiment. Input bacterial community relative
abundance data at the OTU level from the baseline, post-clindamycin, and
post-infection timepoints was used to generate 6 classification models
that predicted \emph{C. difficile} colonization status 7 days
post-infection. The L2-regularized logistic regression models were
trained and tested using the caret package (79) in R as previously
described (80) with the exception that we used 60\% training and 40\%
testing data splits for the cross-validation of the training data to
select the best cost hyperparameter and the testing of the held out test
data to measure model performance. The modified training to testing
ratio was selected to accommodate the small number of samples in the
dataset. Code was modified from
\url{https://github.com/SchlossLab/ML_pipeline_microbiome} to update the
classification outcomes and change the data split ratios. The modified
repository to regenerate this analysis is available at
\url{https://github.com/tomkoset/ML_pipeline_microbiome}.

\textbf{(vii) Statistical analysis.} All statistical tests were
performed in R (v 3.5.2) (78). The Kruskal-Wallis test was used to
analyze differences in \emph{C. difficile} CFU, mouse weight change, and
alpha diversity across sources with a Benjamini-Hochberg correction for
testing multiple timepoints, followed by pairwise Wilcoxon comparisons
with Benjamini-Hochberg correction. For taxonomic analysis and
generation of logistic regression model input data, \emph{C. difficile}
(OTU 20) was removed. Bacterial relative abundances that varied across
sources at the OTU level were identified with the Kruskal-Wallis test
with Benjamini-Hochberg correction for testing all identified OTUs,
followed by pairwise Wilcoxon comparisons with Benjamini-Hochberg
correction. OTUs impacted by clindamycin treatment were identified using
the Wilcoxon signed rank test with matched pairs of mice samples for day
-1 and day 0. To determine whether classification models had better
performance (test AUROCs) than random chance (0.5), we used the
one-sample Wilcoxon signed rank test. To examine whether there was an
overall difference in predictive performance across the 6 classification
models we used the Kruskal-Wallis test followed by pairwise Wilcoxan
comparisons with Benjamini-Hochberg correction for multiple hypothesis
testing. The tidyverse package was used to wrangle and graph data (v
1.3.0) (81).

\textbf{(viii) Code availability.} Code for all data analysis and
generating this manuscript is available at
\url{https://github.com/SchlossLab/Tomkovich_Vendor_XXXX_2020}.

\textbf{(ix) Data availability.} The 16S rRNA sequencing data have been
deposited in the National Center for Biotechnology Information Sequence
Read Archive (BioProject Accession no. PRJNA608529).

\newpage

\subsection{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Teng2019}{}
1. Teng C, Reveles KR, Obodozie-Ofoegbu OO, Frei CR. 2019.
\emph{Clostridium difficile} infection risk with important antibiotic
classes: An analysis of the FDA adverse event reporting system.
International Journal of Medical Sciences 16:630--635.

\hypertarget{ref-Kelly2012}{}
2. Kelly C. 2012. Can we identify patients at high risk of recurrent
\emph{Clostridium difficile} infection? Clinical Microbiology and
Infection 18:21--27.

\hypertarget{ref-Zacharioudakis2015}{}
3. Zacharioudakis IM, Zervou FN, Pliakos EE, Ziakas PD, Mylonakis E.
2015. Colonization with toxinogenic \emph{C. difficile} upon hospital
admission, and risk of infection: A systematic review and meta-analysis.
American Journal of Gastroenterology 110:381--390.

\hypertarget{ref-Crobach2018}{}
4. Crobach MJT, Vernon JJ, Loo VG, Kong LY, Péchiné S, Wilcox MH,
Kuijper EJ. 2018. Understanding \emph{Clostridium difficile}
colonization. Clinical Microbiology Reviews 31.

\hypertarget{ref-Zhang2015}{}
5. Zhang L, Dong D, Jiang C, Li Z, Wang X, Peng Y. 2015. Insight into
alteration of gut microbiota in \emph{Clostridium difficile} infection
and asymptomatic c. difficile colonization. Anaerobe 34:1--7.

\hypertarget{ref-VanInsberghe2020}{}
6. VanInsberghe D, Elsherbini JA, Varian B, Poutahidis T, Erdman S, Polz
MF. 2020. Diarrhoeal events can trigger long-term \emph{Clostridium
difficile} colonization with recurrent blooms. Nature Microbiology
5:642--650.

\hypertarget{ref-Mancabelli2017}{}
7. Mancabelli L, Milani C, Lugli GA, Turroni F, Cocconi D, Sinderen D
van, Ventura M. 2017. Identification of universal gut microbial
biomarkers of common human intestinal diseases by meta-analysis. FEMS
Microbiology Ecology 93.

\hypertarget{ref-Duvallet2017}{}
8. Duvallet C, Gibbons SM, Gurry T, Irizarry RA, Alm EJ. 2017.
Meta-analysis of gut microbiome studies identifies disease-specific and
shared responses. Nature Communications 8.

\hypertarget{ref-Seekatz2016}{}
9. Seekatz AM, Rao K, Santhosh K, Young VB. 2016. Dynamics of the fecal
microbiome in patients with recurrent and nonrecurrent \emph{Clostridium
difficile} infection. Genome Medicine 8.

\hypertarget{ref-Khanna2016}{}
10. Khanna S, Montassier E, Schmidt B, Patel R, Knights D, Pardi DS,
Kashyap PC. 2016. Gut microbiome predictors of treatment response and
recurrence in primary \emph{Clostridium difficile} infection. Alimentary
Pharmacology \& Therapeutics 44:715--727.

\hypertarget{ref-Pakpour2017}{}
11. Pakpour S, Bhanvadia A, Zhu R, Amarnani A, Gibbons SM, Gurry T, Alm
EJ, Martello LA. 2017. Identifying predictive features of
\emph{Clostridium difficile} infection recurrence before, during, and
after primary antibiotic treatment. Microbiome 5.

\hypertarget{ref-Lee2020}{}
12. Lee AA, Rao K, Limsrivilai J, Gillilland M, Malamet B, Briggs E,
Young VB, Higgins PDR. 2020. Temporal gut microbial changes predict
recurrent \emph{Clostridioides difficile} infection in patients with and
without ulcerative colitis. Inflammatory Bowel Diseases.

\hypertarget{ref-Hutton2014}{}
13. Hutton ML, Mackin KE, Chakravorty A, Lyras D. 2014. Small animal
models for the study of \emph{Clostridium difficile} disease
pathogenesis. FEMS Microbiology Letters 352:140--149.

\hypertarget{ref-Chen2008}{}
14. Chen X, Katchar K, Goldsmith JD, Nanthakumar N, Cheknis A, Gerding
DN, Kelly CP. 2008. A mouse model of \emph{Clostridium
difficile}-associated disease. Gastroenterology 135:1984--1992.

\hypertarget{ref-Best2012}{}
15. Best EL, Freeman J, Wilcox MH. 2012. Models for the study of
\emph{Clostridium difficile} infection. Gut Microbes 3:145--167.

\hypertarget{ref-Baxter2014}{}
16. Baxter NT, Wan JJ, Schubert AM, Jenior ML, Myers P, Schloss PD.
2014. Intra- and interindividual variations mask interspecies variation
in the microbiota of sympatric peromyscus populations. Applied and
Environmental Microbiology 81:396--404.

\hypertarget{ref-Nagpal2018}{}
17. Nagpal R, Wang S, Woods LCS, Seshie O, Chung ST, Shively CA,
Register TC, Craft S, McClain DA, Yadav H. 2018. Comparative microbiome
signatures and short-chain fatty acids in mouse, rat, non-human primate,
and human feces. Frontiers in Microbiology 9.

\hypertarget{ref-Reeves2011}{}
18. Reeves AE, Theriot CM, Bergin IL, Huffnagle GB, Schloss PD, Young
VB. 2011. The interplay between microbiome dynamics and pathogen
dynamics in a murine model of \emph{Clostridium difficile} infection
2:145--158.

\hypertarget{ref-Schubert2015}{}
19. Schubert AM, Sinani H, Schloss PD. 2015. Antibiotic-induced
alterations of the murine gut microbiota and subsequent effects on
colonization resistance against \emph{Clostridium difficile}. mBio 6.

\hypertarget{ref-Jenior2017}{}
20. Jenior ML, Leslie JL, Young VB, Schloss PD. 2017. \emph{Clostridium
difficile} colonizes alternative nutrient niches during infection across
distinct murine gut microbiomes. mSystems 2.

\hypertarget{ref-Jenior2018}{}
21. Jenior ML, Leslie JL, Young VB, Schloss PD. 2018. \emph{Clostridium
difficile} alters the structure and metabolism of distinct cecal
microbiomes during initial infection to promote sustained colonization.
mSphere 3.

\hypertarget{ref-Velazquez2019}{}
22. Velazquez EM, Nguyen H, Heasley KT, Saechao CH, Gil LM, Rogers AWL,
Miller BM, Rolston MR, Lopez CA, Litvak Y, Liou MJ, Faber F, Bronner DN,
Tiffany CR, Byndloss MX, Byndloss AJ, Bäumler AJ. 2019. Endogenous
Enterobacteriaceae underlie variation in susceptibility to
\emph{Salmonella} infection. Nature Microbiology 4:1057--1064.

\hypertarget{ref-Osbelt2020}{}
23. Osbelt L, Thiemann S, Smit N, Lesker TR, Schröter M, Gálvez EJC,
Schmidt-Hohagen K, Pils MC, Mühlen S, Dersch P, Hiller K, Schlüter D,
Neumann-Schaal M, Strowig T. 2020. Variations in microbiota composition
of laboratory mice influence \emph{Citrobacter rodentium} infection via
variable short-chain fatty acid production. PLOS Pathogens 16:e1008448.

\hypertarget{ref-Stough2016}{}
24. Stough JMA, Dearth SP, Denny JE, LeCleir GR, Schmidt NW, Campagna
SR, Wilhelm SW. 2016. Functional characteristics of the gut microbiome
in C57BL/6 mice differentially susceptible to \emph{Plasmodium yoelii}.
Frontiers in Microbiology 7.

\hypertarget{ref-Alegre2019}{}
25. Alegre M-L. 2019. Mouse microbiomes: Overlooked culprits of
experimental variability. Genome Biology 20.

\hypertarget{ref-EtienneMesmin2017}{}
26. Etienne-Mesmin L, Chassaing B, Adekunle O, Mattei LM, Bushman FD,
Gewirtz AT. 2017. Toxin-positive \emph{Clostridium difficile} latently
infect mouse colonies and protect against highly pathogenic \emph{C.
difficile}. Gut 67:860--871.

\hypertarget{ref-Lai2020}{}
27. Lai NY, Musser MA, Pinho-Ribeiro FA, Baral P, Jacobson A, Ma P,
Potts DE, Chen Z, Paik D, Soualhi S, Yan Y, Misra A, Goldstein K,
Lagomarsino VN, Nordstrom A, Sivanathan KN, Wallrapp A, Kuchroo VK,
Nowarski R, Starnbach MN, Shi H, Surana NK, An D, Wu C, Huh JR, Rao M,
Chiu IM. 2020. Gut-innervating nociceptor neurons regulate peyer's patch
microfold cells and SFB levels to mediate \emph{Salmonella} host
defense. Cell 180:33--49.e22.

\hypertarget{ref-Thiemann2017}{}
28. Thiemann S, Smit N, Roy U, Lesker TR, Gálvez EJ, Helmecke J, Basic
M, Bleich A, Goodman AL, Kalinke U, Flavell RA, Erhardt M, Strowig T.
2017. Enhancement of IFNgamma production by distinct commensals
ameliorates \emph{Salmonella}-induced disease. Cell Host \& Microbe
21:682--694.e5.

\hypertarget{ref-Rolig2013}{}
29. Rolig AS, Cech C, Ahler E, Carter JE, Ottemann KM. 2013. The degree
of \emph{Helicobacter pylori}-triggered inflammation is manipulated by
preinfection host microbiota. Infection and Immunity 81:1382--1389.

\hypertarget{ref-Ge2018}{}
30. Ge Z, Sheh A, Feng Y, Muthupalani S, Ge L, Wang C, Kurnick S,
Mannion A, Whary MT, Fox JG. 2018. \emph{Helicobacter pylori}-infected
C57BL/6 mice with different gastrointestinal microbiota have contrasting
gastric pathology, microbial and host immune responses. Scientific
Reports 8.

\hypertarget{ref-Lawley2013}{}
31. Lawley TD, Young VB. 2013. Murine models to study \emph{Clostridium
difficile} infection and transmission. Anaerobe 24:94--97.

\hypertarget{ref-Buffie2011}{}
32. Buffie CG, Jarchum I, Equinda M, Lipuma L, Gobourne A, Viale A,
Ubeda C, Xavier J, Pamer EG. 2011. Profound alterations of intestinal
microbiota following a single dose of clindamycin results in sustained
susceptibility to \emph{Clostridium difficile}-induced colitis.
Infection and Immunity 80:62--73.

\hypertarget{ref-Buffie2014}{}
33. Buffie CG, Bucci V, Stein RR, McKenney PT, Ling L, Gobourne A, No D,
Liu H, Kinnebrew M, Viale A, Littmann E, Brink MRM van den, Jenq RR,
Taur Y, Sander C, Cross JR, Toussaint NC, Xavier JB, Pamer EG. 2014.
Precision microbiome reconstitution restores bile acid mediated
resistance to \emph{Clostridium difficile}. Nature 517:205--208.

\hypertarget{ref-Spinler2016}{}
34. Spinler JK, Brown A, Ross CL, Boonma P, Conner ME, Savidge TC. 2016.
Administration of probiotic kefir to mice with \emph{Clostridium
difficile} infection exacerbates disease. Anaerobe 40:54--57.

\hypertarget{ref-Markey2018}{}
35. Markey L, Shaban L, Green ER, Lemon KP, Mecsas J, Kumamoto CA. 2018.
Pre-colonization with the commensal fungus candida albicans reduces
murine susceptibility to \emph{Clostridium difficile} infection. Gut
Microbes 1--13.

\hypertarget{ref-McKee2018}{}
36. McKee RW, Aleksanyan N, Garrett EM, Tamayo R. 2018. Type IV pili
promote \emph{Clostridium difficile} adherence and persistence in a
mouse model of infection. Infection and Immunity 86.

\hypertarget{ref-Yamaguchi2020}{}
37. Yamaguchi T, Konishi H, Aoki K, Ishii Y, Chono K, Tateda K. 2020.
The gut microbiome diversity of \emph{Clostridioides
difficile}-inoculated mice treated with vancomycin and fidaxomicin.
Journal of Infection and Chemotherapy 26:483--491.

\hypertarget{ref-Stroke2018}{}
38. Stroke IL, Letourneau JJ, Miller TE, Xu Y, Pechik I, Savoly DR, Ma
L, Sturzenbecker LJ, Sabalski J, Stein PD, Webb ML, Hilbert DW. 2018.
Treatment of \emph{Clostridium difficile} infection with a
small-molecule inhibitor of toxin UDP-glucose hydrolysis activity.
Antimicrobial Agents and Chemotherapy 62.

\hypertarget{ref-Quigley2019}{}
39. Quigley L, Coakley M, Alemayehu D, Rea MC, Casey PG, O'Sullivan,
Murphy E, Kiely B, Cotter PD, Hill C, Ross RP. 2019. \emph{Lactobacillus
gasseri} APC 678 reduces shedding of the pathogen \emph{Clostridium
difficile} in a murine model. Frontiers in Microbiology 10.

\hypertarget{ref-Mullish2019}{}
40. Mullish BH, McDonald JAK, Pechlivanis A, Allegretti JR, Kao D,
Barker GF, Kapila D, Petrof EO, Joyce SA, Gahan CGM, Glegola-Madejska I,
Williams HRT, Holmes E, Clarke TB, Thursz MR, Marchesi JR. 2019.
Microbial bile salt hydrolases mediate the efficacy of faecal microbiota
transplant in the treatment of recurrent \emph{Clostridioides difficile}
infection. Gut 68:1791--1800.

\hypertarget{ref-Ross2016}{}
41. Ross CL, Spinler JK, Savidge TC. 2016. Structural and functional
changes within the gut microbiota and susceptibility to
\emph{Clostridium difficile} infection. Anaerobe 41:37--43.

\hypertarget{ref-Nguyen2015}{}
42. Nguyen TLA, Vieira-Silva S, Liston A, Raes J. 2015. How informative
is the mouse for human gut microbiota research? Disease Models \&
Mechanisms 8:1--16.

\hypertarget{ref-Tomkovich2019}{}
43. Tomkovich S, Lesniak NA, Li Y, Bishop L, Fitzgerald MJ, Schloss PD.
2019. The proton pump inhibitor omeprazole does not promote
\emph{Clostridioides difficile} colonization in a murine model. mSphere
4.

\hypertarget{ref-Guh2018}{}
44. Guh AY, Kutty PK. 2018. \emph{Clostridioides difficile} infection
169:ITC49.

\hypertarget{ref-Theriot2011}{}
45. Theriot CM, Koumpouras CC, Carlson PE, Bergin II, Aronoff DM, Young
VB. 2011. Cefoperazone-treated mice as an experimental platform to
assess differential virulence of \emph{Clostridium difficile} strains.
Gut Microbes 2:326--334.

\hypertarget{ref-Lawley2009}{}
46. Lawley TD, Clare S, Walker AW, Goulding D, Stabler RA, Croucher N,
Mastroeni P, Scott P, Raisen C, Mottram L, Fairweather NF, Wren BW,
Parkhill J, Dougan G. 2009. Antibiotic treatment of \emph{Clostridium
difficile} carrier mice triggers a supershedder state, spore-mediated
transmission, and severe disease in immunocompromised hosts. Infection
and Immunity 77:3661--3669.

\hypertarget{ref-Lawley2012}{}
47. Lawley TD, Clare S, Walker AW, Stares MD, Connor TR, Raisen C,
Goulding D, Rad R, Schreiber F, Brandt C, Deakin LJ, Pickard DJ, Duncan
SH, Flint HJ, Clark TG, Parkhill J, Dougan G. 2012. Targeted restoration
of the intestinal microbiota with a simple, defined bacteriotherapy
resolves relapsing \emph{Clostridium difficile} disease in mice. PLoS
Pathogens 8:e1002995.

\hypertarget{ref-Jump2014}{}
48. Jump RLP, Polinkovsky A, Hurless K, Sitzlar B, Eckart K, Tomas M,
Deshpande A, Nerandzic MM, Donskey CJ. 2014. Metabolomics analysis
identifies intestinal microbiota-derived biomarkers of colonization
resistance in clindamycin-treated mice. PLoS ONE 9:e101267.

\hypertarget{ref-NagaoKitamoto2020}{}
49. Nagao-Kitamoto H, Leslie JL, Kitamoto S, Jin C, Thomsson KA,
Gillilland MG, Kuffa P, Goto Y, Jenq RR, Ishii C, Hirayama A, Seekatz
AM, Martens EC, Eaton KA, Kao JY, Fukuda S, Higgins PDR, Karlsson NG,
Young VB, Kamada N. 2020. Interleukin-22-mediated host glycosylation
prevents \emph{Clostridioides difficile} infection by modulating the
metabolic activity of the gut microbiota. Nature Medicine 26:608--617.

\hypertarget{ref-Battaglioli2018}{}
50. Battaglioli EJ, Hale VL, Chen J, Jeraldo P, Ruiz-Mojica C, Schmidt
BA, Rekdal VM, Till LM, Huq L, Smits SA, Moor WJ, Jones-Hall Y, Smyrk T,
Khanna S, Pardi DS, Grover M, Patel R, Chia N, Nelson H, Sonnenburg JL,
Farrugia G, Kashyap PC. 2018. \emph{Clostridioides difficile} uses amino
acids associated with gut microbial dysbiosis in a subset of patients
with diarrhea. Science Translational Medicine 10:eaam7019.

\hypertarget{ref-Robinson2014}{}
51. Robinson CD, Auchtung JM, Collins J, Britton RA. 2014. Epidemic
\emph{Clostridium difficile} strains demonstrate increased competitive
fitness compared to nonepidemic isolates. Infection and Immunity
82:2815--2825.

\hypertarget{ref-Collins2015}{}
52. Collins J, Auchtung JM, Schaefer L, Eaton KA, Britton RA. 2015.
Humanized microbiota mice as a model of recurrent \emph{Clostridium
difficile} disease. Microbiome 3.

\hypertarget{ref-Collins2018}{}
53. Collins J, Robinson C, Danhof H, Knetsch CW, Leeuwen HC van, Lawley
TD, Auchtung JM, Britton RA. 2018. Dietary trehalose enhances virulence
of epidemic \emph{Clostridium difficile}. Nature 553:291--294.

\hypertarget{ref-Hryckowian2018}{}
54. Hryckowian AJ, Treuren WV, Smits SA, Davis NM, Gardner JO, Bouley
DM, Sonnenburg JL. 2018. Microbiota-accessible carbohydrates suppress
\emph{Clostridium difficile} infection in a murine model. Nature
Microbiology 3:662--669.

\hypertarget{ref-Fouladi2020}{}
55. Fouladi F, Glenny EM, Bulik-Sullivan EC, Tsilimigras MCB, Sioda M,
Thomas SA, Wang Y, Djukic Z, Tang Q, Tarantino LM, Bulik CM, Fodor AA,
Carroll IM. 2020. Sequence variant analysis reveals poor correlations in
microbial taxonomic abundance between humans and mice after gnotobiotic
transfer. The ISME Journal.

\hypertarget{ref-Walter2020}{}
56. Walter J, Armet AM, Finlay BB, Shanahan F. 2020. Establishing or
exaggerating causality for the gut microbiome: Lessons from human
microbiota-associated rodents. Cell 180:221--232.

\hypertarget{ref-Rasmussen2019}{}
57. Rasmussen TS, Vries L de, Kot W, Hansen LH, Castro-Mejía JL,
Vogensen FK, Hansen AK, Nielsen DS. 2019. Mouse vendor influence on the
bacterial and viral gut composition exceeds the effect of diet. Viruses
11:435.

\hypertarget{ref-Mims2020}{}
58. Mims TS, Abdallah QA, Watts S, White C, Han J, Willis KA, Pierre JF.
2020. Variability in interkingdom gut microbiomes between different
commercial vendors shapes fat gain in response to diet. The FASEB
Journal 34:1--1.

\hypertarget{ref-Stewart2019}{}
59. Stewart DB, Wright JR, Fowler M, McLimans CJ, Tokarev V, Amaniera I,
Baker O, Wong H-T, Brabec J, Drucker R, Lamendella R. 2019. Integrated
meta-omics reveals a fungus-associated bacteriome and distinct
functional pathways in \emph{Clostridioides difficile} infection.
mSphere 4.

\hypertarget{ref-Ott2017}{}
60. Ott SJ, Waetzig GH, Rehman A, Moltzau-Anderson J, Bharti R, Grasis
JA, Cassidy L, Tholey A, Fickenscher H, Seegert D, Rosenstiel P,
Schreiber S. 2017. Efficacy of sterile fecal filtrate transfer for
treating patients with \emph{Clostridium difficile} infection.
Gastroenterology 152:799--811.e7.

\hypertarget{ref-Zuo2017}{}
61. Zuo T, Wong SH, Lam K, Lui R, Cheung K, Tang W, Ching JYL, Chan PKS,
Chan MCW, Wu JCY, Chan FKL, Yu J, Sung JJY, Ng SC. 2017. Bacteriophage
transfer during faecal microbiota transplantation in \emph{Clostridium
difficile} infection is associated with treatment outcome. Gut
gutjnl--2017--313952.

\hypertarget{ref-Zuo2018}{}
62. Zuo T, Wong SH, Cheung CP, Lam K, Lui R, Cheung K, Zhang F, Tang W,
Ching JYL, Wu JCY, Chan PKS, Sung JJY, Yu J, Chan FKL, Ng SC. 2018. Gut
fungal dysbiosis correlates with reduced efficacy of fecal microbiota
transplantation in \emph{Clostridium difficile} infection. Nature
Communications 9.

\hypertarget{ref-Robinson2019}{}
63. Robinson JI, Weir WH, Crowley JR, Hink T, Reske KA, Kwon JH, Burnham
C-AD, Dubberke ER, Mucha PJ, Henderson JP. 2019. Metabolomic networks
connect host-microbiome processes to human \emph{Clostridioides
difficile} infections. Journal of Clinical Investigation 129:3792--3806.

\hypertarget{ref-Fletcher2018}{}
64. Fletcher JR, Erwin S, Lanzas C, Theriot CM. 2018. Shifts in the gut
metabolome and \emph{Clostridium difficile} transcriptome throughout
colonization and infection in a mouse model. mSphere 3.

\hypertarget{ref-Xiao2015}{}
65. Xiao L, Feng Q, Liang S, Sonne SB, Xia Z, Qiu X, Li X, Long H, Zhang
J, Zhang D, Liu C, Fang Z, Chou J, Glanville J, Hao Q, Kotowska D,
Colding C, Licht TR, Wu D, Yu J, Sung JJY, Liang Q, Li J, Jia H, Lan Z,
Tremaroli V, Dworzynski P, Nielsen HB, Bäckhed F, Doré J, Chatelier EL,
Ehrlich SD, Lin JC, Arumugam M, Wang J, Madsen L, Kristiansen K. 2015. A
catalog of the mouse gut metagenome. Nature Biotechnology 33:1103--1108.

\hypertarget{ref-Vital2019}{}
66. Vital M, Rud T, Rath S, Pieper DH, Schlüter D. 2019. Diversity of
bacteria exhibiting bile acid-inducible 7alpha-dehydroxylation genes in
the human gut. Computational and Structural Biotechnology Journal
17:1016--1019.

\hypertarget{ref-Fransen2015}{}
67. Fransen F, Zagato E, Mazzini E, Fosso B, Manzari C, Aidy SE,
Chiavelli A, D'Erchia AM, Sethi MK, Pabst O, Marzano M, Moretti S,
Romani L, Penna G, Pesole G, Rescigno M. 2015. BALB/c and C57BL/6 mice
differ in polyreactive IgA abundance, which impacts the generation of
antigen-specific IgA and microbiota diversity. Immunity 43:527--540.

\hypertarget{ref-Ivanov2009}{}
68. Ivanov II, Atarashi K, Manel N, Brodie EL, Shima T, Karaoz U, Wei D,
Goldfarb KC, Santee CA, Lynch SV, Tanoue T, Imaoka A, Itoh K, Takeda K,
Umesaki Y, Honda K, Littman DR. 2009. Induction of intestinal th17 cells
by segmented filamentous bacteria. Cell 139:485--498.

\hypertarget{ref-Azrad2018}{}
69. Azrad M, Hamo Z, Tkhawkho L, Peretz A. 2018. Elevated serum
immunoglobulin a levels in patients with \emph{Clostridium difficile}
infection are associated with mortality. Pathogens and Disease 76.

\hypertarget{ref-Saleh2019}{}
70. Saleh MM, Frisbee AL, Leslie JL, Buonomo EL, Cowardin CA, Ma JZ,
Simpson ME, Scully KW, Abhyankar MM, Petri WA. 2019. Colitis-induced
th17 cells increase the risk for severe subsequent \emph{Clostridium
difficile} infection. Cell Host \& Microbe 25:756--765.e5.

\hypertarget{ref-Kozich2013}{}
71. Kozich JJ, Westcott SL, Baxter NT, Highlander SK, Schloss PD. 2013.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq illumina sequencing
platform. Applied and Environmental Microbiology 79:5112--5120.

\hypertarget{ref-Sze2019}{}
72. Sze MA, Schloss PD. 2019. The impact of DNA polymerase and number of
rounds of amplification in PCR on 16S rRNA gene sequence data. mSphere
4.

\hypertarget{ref-Schloss2009}{}
73. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, Lesniewski RA, Oakley BB, Parks DH, Robinson CJ, Sahl JW, Stres B,
Thallinger GG, Horn DJV, Weber CF. 2009. Introducing mothur:
Open-source, platform-independent, community-supported software for
describing and comparing microbial communities. Applied and
Environmental Microbiology 75:7537--7541.

\hypertarget{ref-Quast2012}{}
74. Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies
J, Glöckner FO. 2012. The SILVA ribosomal RNA gene database project:
Improved data processing and web-based tools. Nucleic Acids Research
41:D590--D596.

\hypertarget{ref-Cole2013}{}
75. Cole JR, Wang Q, Fish JA, Chai B, McGarrell DM, Sun Y, Brown CT,
Porras-Alfaro A, Kuske CR, Tiedje JM. 2013. Ribosomal database project:
Data and tools for high throughput rRNA analysis. Nucleic Acids Research
42:D633--D642.

\hypertarget{ref-Westcott2017}{}
76. Westcott SL, Schloss PD. 2017. OptiClust, an improved method for
assigning amplicon-based sequence data to operational taxonomic units.
mSphere 2.

\hypertarget{ref-Vegan2018}{}
77. Oksanen J, Blanchet FG, Friendly M, Kindt R, Legendre P, McGlinn D,
Minchin PR, O'Hara RB, Simpson GL, Solymos P, Stevens MHH, Szoecs E,
Wagner H. 2018. Vegan: Community ecology package.

\hypertarget{ref-r_citation_2018}{}
78. R Core Team. 2018. R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria.

\hypertarget{ref-Kuhn2008}{}
79. Kuhn M. 2008. Building predictive models inRUsing thecaretPackage.
Journal of Statistical Software 28.

\hypertarget{ref-Topcuoglu2020}{}
80. Topçuoğlu BD, Lesniak NA, Ruffin MT, Wiens J, Schloss PD. 2020. A
framework for effective application of machine learning to
microbiome-based classification problems. mBio 11.

\hypertarget{ref-Tidyverse2019}{}
81. Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R,
Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E,
Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K,
Vaughan D, Wilke C, Woo K, Yutani H. 2019. Welcome to the tidyverse.
Journal of Open Source Software 4:1686.

\newpage

\subsection{Figures}\label{figures}

\includegraphics{figure_1.pdf} \textbf{Figure 1. Microbiota variation is
high between mice from different sources.} A-B. Number of observed OTUs
(A) and Shannon diversity index values (B) across sources of mice at
baseline (day -1) of the experiment. Differences across sources were
analyzed by Kruskal-Wallis test with Benjamini-Hochberg correction for
testing each day of the experiment and the adjusted \emph{P} value was
\textless{} 0.05 for panel A (Table S1). None of the \emph{P} values
from pairwise Wilcoxon comparisons between sources were significant
after Benjamini-Hochberg correction (Table S1). Gray lines represent the
median values for each source of mice. C. Principal Coordinates Analysis
(PCoA) of \(\theta_{YC}\) distances of baseline stool samples.
Permutational multivariate analysis of variance (PERMANOVA) analysis
demonstrated that source and the interaction between source and cage
explained most of the variation (combined R\textsuperscript{2} = 0.90,
\emph{P} = 0.0001, see Table S2). For A-C: each symbol representing the
value for a stool sample from an individual mouse, circles represent
experiment 1 mice and triangles represent experiment 2 mice. D. Plots
highlighting the median (point) and interquantile range (colored lines)
of the relative abundances for the top 20 bacteria out of the 268 OTUs
that varied across sources at baseline (Table S3).

\newpage

\includegraphics{figure_2.pdf}

\textbf{Figure 2. Clindamycin is sufficient to promote \emph{C.
difficile} colonization in all mice, but clearance time varies across
sources of C57BL/6 mice.} A. Setup of the experimental timeline. Mice
for the experiments were obtained from 6 different sources: the Schloss
(N = 8) and Young lab (N = 9) colonies at the University of Michigan,
the Jackson Laboratory (N = 8), Charles River Laboratory (N = 8),
Taconic Biosciences (N = 8), and Envigo (N = 8). All mice were
administered 10 mg/kg clindamycin intraperitoneally (IP) 1 day before
challenge with \emph{C. difficile} 630 spores on day 0. Mice were
weighed and feces was collected daily through the end of the experiment
(9 days post-infection). Note: 3 mice died during course of experiment.
1 Taconic mouse prior to infection and 1 Jackson and 1 Envigo mouse
between 1- and 3-days post-infection. B. \emph{C. difficile} CFU/gram
stool measured over time (N = 20-49 mice per timepoint) via serial
dilutions. The black line represents the limit of detection for the
first serial dilution. CFU quantification data was not available for
each mouse due to early deaths, stool sampling difficulties, and not
plating all of the serial dilutions. C. Mouse weight change measured in
grams over time (N = 45-49 mice per timepoint), all mice were normalized
to the weight recorded 1 day before infection. For B-C: timepoints where
differences across sources of mice were statistically significant by
Kruskal-Wallis test with Benjamini-Hochberg correction for testing
across multiple days (Table S4 and Table S5) are reflected by the
asterisk(s) above each timepoint (*, \emph{P} \textless{} 0.05). Lines
represent the median for each source and circles represent individual
mice from experiment 1 while triangles represent mice from experiment 2.

\newpage

\includegraphics{figure_3.pdf} \textbf{Figure 3. Clindamycin treatment
alters bacteria in all sources, but a subset of bacterial differences
across sources persists.} A-B. Number of observed OTUs (A) and Shannon
diversity index values (B) across sources of mice after clindamycin
treatment (day 0). Data were analyzed by Kruskal-Wallis test with
Benjamini-Hochberg correction for testing each day of the experiment and
the adjusted \emph{P} value was \textless{} 0.05. Significant \emph{P}
values from the pairwise Wilcoxon comparisons between sources with
Benjamini-Hochberg correction are shown (Table S1). C. PCoA of
\(\theta_{YC}\) distances from stools collected post-clindamycin.
PERMANOVA analysis demonstrated that source and the interaction between
source and cage explained most of the variation observed
post-clindamycin (combined R\textsuperscript{2} = 0.99, \emph{P} =
0.0001, see Table S2).For A-C, each symbol represents a stool sample
from an individual mouse, with circles representing experiment 1 mice
and triangles representing experiment 2 mice. D. Plots highlighting the
median (point) and interquantile range (colored lines) of the top 10
OTUs out of 153 with relative abundances that changed after clindamycin
treatment (adjusted \emph{P} value \textless{} 0.05). Data were analyzed
by Wilcoxon signed rank test limited to mice that had paired sequence
data for day -1 and 0 (N = 31), with Benjamini-Hochberg correction for
testing all identified OTUs (Table S6). The gray vertical line indicates
the limit of detection. E. Plots highlighting the median (point) and
interquantile range (colored lines) of the relative abundances for the
18 OTUs that varied across sources after clindamycin treatment (Table
S3).

\newpage

\includegraphics{figure_4.pdf} \textbf{Figure 4. Microbiota variation
across sources is maintained after \emph{C. difficile} challenge.} A-B.
Number of observed OTUs (A) and Shannon diversity index values (B)
across sources of mice post-infection (day 1). Data were analyzed by
Kruskal-Wallis test with Benjamini-Hochberg correction for testing each
day of the experiment and the adjusted \emph{P} value was \textless{}
0.05. Significant \emph{P} values from the pairwise Wilcoxon comparisons
between sources with Benjamini-Hochberg correction are shown (Table S1).
PCoA of \(\theta_{YC}\) distances of post-infection stool samples.
PERMANOVA analysis demonstrated that source and the interaction between
source and cage explained most of the variation (combined
R\textsuperscript{2} = 0.88, \emph{P} = 0.0001, Table S2). For A-C: each
symbol representing the value for a stool sample from an individual
mouse, circles represent experiment 1 mice and triangles represent
experiment 2 mice. D. Plots highlighting the median (point) and
interquantile range (colored lines) of the relative abundances for the
top 20 bacteria out of the 44 OTUs that varied across sources
post-infection. The gray vertical line indicates the limit of detection.
For each timepoint OTUs with differential relative abundances across
sources of mice were identified by Kruskal-Wallis test with
Benjamini-Hochberg correction for testing all identified OTUs (Table
S3).

\newpage

\includegraphics{figure_5.pdf} \textbf{Figure 5. Bacteria that influence
whether mice cleared \emph{C. difficile} by day 7.} A. Baseline relative
abundance data for 3 of the OTUs from the classification model based on
day 0 OTU relative abundances that significantly varied across sources
of mice and had high relative abundances in the community. Symbols
represent the relative abundance data for an individual mouse, circles
represent mice that cleared \emph{C. difficile} by day 7, X-shapes
represent mice that were still colonized with \emph{C. difficile}, and
open circles represent mice that did not have \emph{C. difficile} CFU
counts for day 7 post-infection. Gray lines indicate the median relative
abundances for each source. Asterisks are shown for pairwise Wilcoxon
comparisons with Benjamini-Hochberg correction where \emph{P}
\textless{} 0.05. B. Venn diagram that combines Fig. S4 summaries of
OTUs that were important to the day -1, 0, and 1 classification models
(Table S8) and either overlapped with taxa that varied across sources at
the same timepoint, were impacted by clindamycin treatment, or both. See
Fig. S4 for separate comparisons of taxa from the day -1, 0, and 1
classification models. Red OTUs were important to more than 1
classification model.

\newpage

\includegraphics{figure_6.pdf} \textbf{Figure 6: Key OTUs vary across
sources throughout the experiment.} A-D. Relative abundances of red OTUs
from Fig. 5A that were important for at least two classification models
are shown over time. A. \emph{Bacteroides} (OTU 2), which varied across
sources throughout the experiment. B-C. \emph{Enterobacteriaceae} (B)
and \emph{Enterococcus} (C), which significantly varied across sources
and were impacted by clindamycin treatment. D. \emph{Porphyromonadaceae}
(OTU 7), which was significantly impacted by clindamycin treatment and
examining relative abundance dynamics over the course of the experiment,
revealed timepoints where relative abundances also significantly varied
across sources of mice. Symbols represent the relative abundance data
for an individual mouse, circles represent mice that cleared \emph{C.
difficile} by day 7, X-shapes represent mice that were still colonized
with \emph{C. difficile}, and open circles represent mice that did not
have \emph{C. difficile} CFU counts for day 7 post-infection. Colored
lines indicate the median relative abundances for each source. The gray
horizontal line represents the limit of detection. Timepoints where
differences across sources of mice were statistically significant by
Kruskal-Wallis test with Benjamini-Hochberg correction for testing
across multiple days (Table S9) are identified by the asterisk(s) above
each timepoint (*, P \textless{} 0.05).

\newpage

\includegraphics{figure_S1.pdf} \textbf{Figure S1. Bacterial communities
vary between experiments for some sources.} A-F. PCoA of \(\theta_{YC}\)
distances for the baseline fecal bacterial communities within each
source of mice. Each symbol represents a stool sample from an individual
mouse with color corresponding to experiment and shape representing cage
mates. PERMANOVA was performed within each group to examine the
contributions of experiment and cage to observed variation. Experiment
number and cage only significantly explained observed variation for mice
from the Schloss (combined R\textsuperscript{2} = 0.99; \emph{P} \(\le\)
0.033) and Young (combined R\textsuperscript{2} = 0.95; \emph{P} \(\le\)
0.029) lab colonies (Table S2). G-H: Boxplots of the \(\theta_{YC}\)
distances of the 6 sources of mice relative to mice within the same
source and experiment (G) or mice within the same source and between
experiments (H) at baseline. Symbols represent individual mouse samples:
circles for experiment 1 and triangles for experiment 2.

\newpage

\includegraphics{figure_S2.pdf} \textbf{Figure S2. \emph{C. difficile}
CFU variation across sources varies slightly between the 2 experiments.}
A-B. \emph{C. difficile} CFU/gram of stool quantification over time for
experiment 1 (A) and 2 (B). Experiments were conducted approximately 3
months apart. Lines represent the median CFU for each source, symbols
represent individual mice and the black line represents the limit of
detection. C. \emph{C. difficile} CFU/gram stool on day 7 post-infection
across sources of mice with asterisks for pairwise Wilcoxon comparisons
with Benjamini-Hochberg correction where \emph{P} \textless{} 0.05. D.
Mouse weight change 2 days post-infection across sources of mice, no
pairwise Wilcoxon comparisons were significant after Benjamini-Hochberg
correction. For C-D: circles represent experiment 1 mice, triangles
represent experiment 2 mice and gray lines indicate the median values
for each group. E. Percent of mice that were colonized with \emph{C.
difficile} over the course of the experiment. Each day the percent is
calculated based on the mice where \emph{C. difficile} CFU was
quantified for that particular day. Total N for each day: day 1 (N =
42), day 2 (N = 20), day 3 (N = 39), day 4 (N = 29), day 5 (N = 43), day
6 (N = 34), day 7 (N = 40), day 8 (N = 36), and day 9 (N = 46).

\newpage

\includegraphics{figure_S3.pdf} \textbf{Figure S3. Bacterial community
composition before, after clindamycin perturbation, and post-infection
can predict \emph{C. difficile} colonization status 7 days
post-challenge.} A. Bar graph visualizations of overall day 7 C.
difficile colonization status that were used as classification outcomes
to build L2-regularized logistic regression models. Mice were classified
as colonized or cleared (not detectable at the limit of detection of 100
CFU) based on CFU g/stool data from 7 days post-infection. B. \emph{C.
difficile} CFU status on Day 7 within each mouse source. N = 5-9 mice
per group. C. L2-regularized logistic regression classification model
area under the receiving operator characteristic curve (AUROCs) to
predict \emph{C. difficile} CFU on day 7 post-infectoin (Fig. 1D, Fig.
S3) based on the OTU community relative abundances at baseline (day -1),
post-clindamycin (day 0), and post-infection (day 1). All models
performed better than random chance (AUROC = 0.5; all \emph{P} \(\le\)
1.6e-17; Table S7) and the model built with post-clindamycin treated
bacterial OTU relative abundances had the best performance
((\emph{P}\textsubscript{FDR} \(\le\) 3.3e-06 for all pairwise
comparisons; Table S7). For list of the 20 OTUs that were ranked as most
important to each model, see Table S8.

\newpage

\includegraphics{figure_S4.pdf} \textbf{Figure S4. OTUs from
classification models based on baseline, post-clindamycin treatment, or
post-infection community data vary by source, clindamycin treatment, or
both.} A-C. Venn diagrams of top 20 important OTUs from baseline (A),
post-clindamycin treatment (B), and post-infection (C) classification
models (Table S8) that overlapped with OTUs that varied across sources
at the corresponding timepoint, were impacted by clindamycin treatment,
or both. Bold OTUs signify OTUs that were important to more than 1
classification model.

\newpage

\subsection{Supplementary Tables and
Movie}\label{supplementary-tables-and-movie}

All supplemental material is available at:
\url{https://github.com/SchlossLab/Tomkovich_Vendor_XXXX_2020/submission}.

\textbf{Movie S1. Large shifts in bacterial community structure occurred
after clindamycin and \emph{C. difficile} infection.} PCoA of
\(\theta_{YC}\) distances animated from 0 through 9 days post-infection.
PERMANOVA analysis indicated source was the variable that explained the
most observed variation across fecal communities (source
R\textsuperscript{2} = 0.35, \emph{P} = 0.0001) followed by interactions
between cage and day of the experiment. Transparency of the circle
corresponds to the day of the experiment, each circle represents a
sample from an individual mouse at a specific timepoint. See Table S2
for PERMANOVA results). Circles represent mice from experiment 1 and
triangles represent mice from expeirment 2.

\textbf{Table S1. Diversity metrics Kruskal-Wallis and pairwise Wilcoxon
statistical results.}

\textbf{Table S2. PERMANOVA results for mice at specific timepoints and
across all timepoints.}

\textbf{Table S3. OTUs with relative abudances that significantly vary
across sources at baseline, post-clindamycin, or post-infection
timepoints.}

\textbf{Table S4. \emph{C. difficile} CFU statistical results.}

\textbf{Table S5. Mouse weight change statistical results.}

\textbf{Table S6. OTUs with relative abudances that significantly
changed after clindamycin treatment.}

\textbf{Table S7. Statistical results of L2-regularized logistic
regression model performances.}

\textbf{Table S8. Top 20 most important OTUs for each of the 3
L2-regularized logistic regression models based on OTU relative
abundance data.}

\textbf{Table S9. OTUs with relative abudances that significantly varied
across sources of mice on at least 1 day of the experiment by
Kruskal-Wallis test.}

\end{document}